# Linear Models

#### CS 4375 - Intro to Machine Learning
#### Dr. Karen Mazidi
#### Author: Leo Nguyen -ldn190002
#### Date: Sep 25, 2022


## Linear Regression Models
- General term
- Strong and weakness


## Physicochemical Properties of Protein Tertiary Structure Data Set
**Citation:** 
ABV - Indian Institute of Information Technology & Management, Gwalior, MP, India.
UCI Machine Learning Repository [http://archive.ics.uci.edu/ml]. Irvine, CA: University of California, School of Information and Computer Science.

**Attribute Information*:**
RMSD-Size of the residue.
F1 - Total surface area.
F2 - Non polar exposed area.
F3 - Fractional area of exposed non polar residue.
F4 - Fractional area of exposed non polar part of residue.
F5 - Molecular mass weighted exposed area.
F6 - Average deviation from standard exposed area of residue.
F7 - Euclidian distance.
F8 - Secondary structure penalty.
F9 - Spacial Distribution constraints (N,K Value).

### Load the data
```{r}
df <- read.csv("data/CASP.csv", header=TRUE)
str(df)
```
### Divide into 80/20 train/test

```{r}
set.seed(1234)
i <- sample(1:nrow(df), 0.8*nrow(df), replace=FALSE)
train <- df[i,]
test <- df[-i,]
```

### Data Exploration
1. Look at the first 10 rows in training data
```{r}
head(train, n=10)
```
2. View the summary of entire training data set
```{r}
summary(train)
```
3. Count number of NA value by column
```{r}
colSums(is.na(train)) 
```
4. Check the correlation of all column in training data
- Based on the number we can see that F1 is highly correlation with F2, F4, F5, F6, F9
```{r}
cor(train)
```
5. Quick visualization on correlation 
- These plot confirm the highly correlation of F1 and other attributes mentioned  above


